\emph{MixFishSim}: highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-01-28 write_up/mixfishsim_ecolmod.tex
Updated text and output latex table of results

\usepackage{lineno,hyperref}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,chains,calc}
\usepackage{hhline}
\usepackage{booktabs}
%\usepackage[textsize=scriptsize]{todonotes}
\usepackage[textsize=scriptsize, disable]{todonotes}  %% to turn off notes
\usepackage{eurosym}

@@

We simulate 50 years of fishing and use the results from the fisheries catch to
draw inference on the underlying population structures. We compare this
inference to i) a simulated fixed-site sampling design commonly used for
fisheries monitoring purposes, and ii) the true underlying population
structures input to the simulation. We use the results to establish the
potential and limitations of fishery-dependent data - an inherently biased
sampling method due to fisher's targeting- to provide a robust picture of
spatiotemporal distributions. Finally, we simulate an area closure based on
areas defined from the known ("real-population") distribution, commercial catch
data and survey data at different temporal and spatial resolutions and assess
their effectiveness on reducing catches of a fish population. \\

We conclude from our simulations that commercial data, while not unbiased,
provides a useful tool for managing catches in mixed fisheries if applied at
the correct spatiotemporal scale. \\

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time with varying knowledge of species distributions
using species-unselective fishing gear. \added[id=PD]{In doing so} fisheries
\deleted[id=PD]{that} catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} overquota catch \added[id=JJ]{when managed by
	single species quotas,} leading to overexploitation of fish populations
\added[id=JJ]{\citep{Ulrich2011a, Batsleer2015}}.  \added[id=JJ]{This
	discarding of fish in excess of quota hampers the ability to limit
	fishing mortality to within sustainable limits \citep{Alverson1994,
		Crowder1998, Rijnsdorp2007}} \deleted[id=PD]{; reducing
	discarding is crucial}and ensure biological and economic sustainability
of fisheries\deleted[id=JJ]{ and implementation of an ecosystem approach to
	fisheries}. As such, there is increasing interest
in technical solutions such as gear and spatial closures as ways of
\replaced[id=PD]{reducing unwanted catch}{avoiding discard\added[id=JJ]{ing of
		fish}}\citep{Kennelly2002, Catchpole2008, Bellido2011}.  \\

\replaced[id=PD]{Changes to spatial fishing patterns have}{Use of spatial
	management as a tools} been proposed as a method to reduce discards
\added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}. However,
\deleted[id=PD]{its} implementation is hampered by lack of knowledge of fish
and fishery spatiotemporal dynamics and understanding of the scale at which
processes are important for management. Understanding the correct scale for
spatial management is crucial in order to implement measures at a resolution
that ensures effective management \citep{Dunn2016} while minimising economic
impact.  For example, a scale that promotes species avoidance for vulnerable or
low quota species while allowing continuance of sustainable fisheries for
available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} an
appropriate scale has been a challenge in the past that has led to ineffectual
measures with unintended consequences such as limited impact towards the
management objective or increased benthic impact on previously unexploited
areas (e.g. the cod closure in the North Sea
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatial information has \added[id=PD]{since} become available through
the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}.  Such
information is, however, patchy and derived from an inherently biased sampling
programme (i.e. targeted fishing). \deleted[id=PD]{ Further, fishers generally only recorded
landings (not catch) on a daily basis\todo{\added[id=JJ]{Its not clear what the
		problem is: landings collected on daily basis or landings
		recorded rather than catch}}.  This leads to questions about
the validity of inference that can be drawn from landings data assigned to VMS
activity pings.}\todo{\added[id=JJ]{This comes as a surprise: I thought this was
		going to be about discards}\added[id=PD]{Agree, have removed
		this to avoid confusion}}\\

@@

\begin{enumerate}
	\item to understand how sampling-derived data reflects the underlying
		population structures. We compare at different spatial and
		temporal aggregations of data the real population to: 
		\begin{enumerate}
			\item the inferred population from a stratified fixed-site sampling survey design
				commonly used for fisheries monitoring purposes, otherwise know as a
				fisheries-independent survey,
			\item the inferred population from our fishery-dependent model which includes
				fishery-induced sampling dynamics.\\
		\end{enumerate}
	\item to understand the impact of data aggregation and source on
		spatial fisheries management measures we simulate a fishery closure to
		protect a species based on different spatial and temporal data
		aggregations:
		\begin{enumerate}
			\item as if the real spatial population structure were
				known,
			\item the fishery-independent inferred population
				structure
			\item the fishery-dependent inferred population
				structure
		\end{enumerate}
		We evaluate the theoretical "benefit" to the population of the
		closure(s), the effect on the other three populations and
		fishery catch.\todo{\added[id=JJ]{If the paper has two goals
				this should be clear from the start, but may be
				better over 2 MSs}\added[id=PD]{I would like to
				keep both parts, but have made clearer in how
				its set out. The closure scenarios form
				validation of the data aggregation, rather than
				effectiveness of the closures themselves - so
				its a continuation of the same question in my
				eyes}}\deleted[id=PD]{.  Further, we extend our
			analysis to a range of spatial and temporal scales to
			assess the impact of these processes on the success of
			the management measure.}\\
\end{enumerate}

@@

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation*}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} * (1 - e^{-(F_{c,d} + M_{c,d})}) *
B_{c,d}
\end{equation*}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. \todo{\added[id=CM]{[link $F$ to
	effort and catchability - as I think we have F as an emergent property
	of the fleets rather than something we solve for (I could be wrong
	though!) - catch for a vessel is a product of catchability and biomass,
	i.e. C = qB, but this catch is summed to solve for F. So its both
	really]}}\\

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha * B_{c,d})}{(\beta + B_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum, $B$ current SSB and
$\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} * e^{(\alpha - \beta * B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases.

@@

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations\replaced[id=JJ]{. An}{,
	with an} advection-diffusion process \deleted[id=JJ]{to}
control\added[id=JJ]{led} \deleted[id=JJ]{how the} population\deleted[id=JJ]{s}
move\replaced[id=JJ]{ment}{d} over time with a moving temperature covariate to
capture temporal dependencies\todo{\added[id=JJ]{What have a temperature
		covariate? Could just use time}\added[id=PD]{Was intended as
		some biological meaning - species thermal tolerances load onto
		the temperature effect - so could be different per species}}.
\deleted[id=JJ]{This was intended to balance realism in population movement,
	capturing the main directed and random processes, and practicality of
	modelling the population rather than individual fish.} \\

For \deleted[id=PD]{the} habitat we define\added[id=PD]{d} a Gaussian random
field process\todo{\added[id=JJ]{Not clear how habitat/GRF affect local
		abundances, only have $B_{y,d}$}\added[id=PD]{Have included
		cell reference, $c$ to make spatial link explicit}}, $\{S(c) :
c \in \mathbb{R}^2\}$, \deleted[id=PD]{that is a stochastic process} where
\added[id=PD]{for} any set of cells $c_{1}, \dots, c_{n}$ \deleted[id=PD]{where
	for each $c_{i} \in \mathbb{R}^2$}, the joint distribution of $S =
\{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian. The distribution is
specified by its \textit{mean function}, $\mu(c) = E[S(c)]$ and its
\textit{covariance function}\todo{\added[id=JM]{Introduce the gamma function,
		and why this covariance structure? Why correlate values in the
		random field?}\added[id=PD]{to allow populations to have
		different aggregation densities: have tried to clarify}},
$\gamma(c,c') = Cov\{S(c),S(c')\}$ \citep{Diggle2007}.\\

The covariance structure affects the smoothness of the surfaces which the
process generates\replaced[id=PD]{;}{, and} we used the \textit{Matérn}
\deleted[id=PD]{family of} covariance structure\deleted[id=PD]{s},
\replaced[id=PD]{as}{one where} the correlation strength weakens the further
the distance apart \deleted[id=PD]{(i.e. the correlation between $S(x)$ and
	$S(x')$ decreases as the distance $u = ||x - x'||$ increases)}.
\added[id=PD]{The Matérn covariance structure models the spatial
	autocorrelation, where animal densities are observed to be more similar
	in nearby locations \citep{Tobler1970,
		F.Dormann2007}}\replaced[id=PD]{. It is}{The \textit{Matérn}
	correlation} is a two-parameter family where: 
\begin{center}
	$\rho(u) = \{2^{\kappa -
		1}\Gamma{\kappa}\}^{-1}(u/\phi)^{\kappa}K_{\kappa}(u/\phi)$
\end{center}
$K_{\kappa}(.)$ is a modified Bessel function of order $\kappa$, $\phi >
0$ is a scale parameter with the dimensions of distance, and $\kappa > 0$,
called the order, is a shape parameter which determines the smoothness of the
underlying process. \\

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}, implementing different parameter settings to affect the
patchiness of the populations. Each population \replaced[id=PD]{was}{is}
initialised at a single location, and subsequently move\replaced[id=PD]{ed}{s}
according to a probabilistic distrib\added[id=PD]{u}tion based on habitat
suitability, temperature and distance from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{e^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$.\\

During specified weeks of the year, the habitat quality
\replaced[id=PD]{was}{is} modified for \added[id=PD]{user-defined} spawning
habitat\deleted[id=PD]{s}, \replaced[id=PD]{resulting in}{meaning} each
population \replaced[id=PD]{having}{has} a concentrated area where spawning takes
place and the population move\replaced[id=PD]{d}{s} towards
\replaced[id=PD]{these cells}{this} in the weeks prior to
spawning.\todo{\added[id=JM]{What does it mean concisely? Areas are assigned?}
\added[id=PD]{Yes, the areas are pre-defined - I have amended to reflect and
	tried to clarify}} \\

The temperature field \replaced[id=PD]{was}{is} simulated to be on a gradient
from a South-Westerly to North-Easterly direction, with temperature in each
cell changing gradually on a week-by-week basis so that initially high
temperature areas cycle\added[id=PD]{d} to lower temperatures and low
temperature areas vice versa. Each population $p$ \replaced[id=PD]{wa}{i}s
assigned a thermal tolerance with mean, $\mu\added[id=PD]{_{p}}$ and variance,
$\sigma^2\added[id=PD]{_{p}}$ so that each cell and population temperature
suitability is defined that:
\begin{equation}
	Tol_{c, p\added[id=PD]{, wk}} = \frac{1}{ \sqrt (2\pi \cdot \sigma^2_{p})} \cdot \exp(-
		\frac{(T_{c\added[id=PD]{, wk}} - \mu_{p})^2}{2 \cdot \sigma^2_{p}} )	
\end{equation}
Where $Tol_{c, p, \added[id=PD]{wk}}$ is the tolerance of population $p$ for
cell $c$ \added[id=PD]{in week $wk$}, $T_{c, \added[id=PD]{wk}}$ is the
temperature in the cell \added[id=PD]{given the week} and
$\mu\added[id=PD]{_{p}}$ and $\sigma^2\added[id=PD]{_{p}}$ the mean and
standard deviation of the population temperature tolerance. \\

\added[id=PD]{The final process resulted in independent populations structure
	and movement patterns, with population movement occuring on a weekly
	basis. This process approximated the demographic shifts in fish
	populations throughout a year with seasonal spawning patterns.}

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determine\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, which
determine\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip.

@@

Each fleet of \textit{n} vessels \replaced[id=PD]{was}{is} characterised by
both a general efficiency, \textit{Q}, and a population specific efficiency,
${Q_{p}}$.  Thus, the product of these parameters affect\replaced[id=PD]{s}{ed}
the overall catch rates for the fleet and the preferential targeting of one
population over another. This, in combination with the parameter choice for the
step-function \added[id=PD]{defined below} (as well as some randomness from the
exploratory fishing process) determine\replaced[id=PD]{d}{s} the preference of
fishing locations for the fleet.  All species prices
\replaced[id=PD]{were}{are} kept the same across fleets \replaced[id=PD]{and
	seasons}{, though can be made to vary seasonally}.

@@

NOTE: THIS IS EXPLORE-EXPLOIT STRATEGY VIZ. BAILEY ET AL POSEIDON MODEL.

Several studies \citep[e.g.][]{Hutton2004, Tidd2012, Girardin2015} have
confirmed past activity and past catch rates are strong predictors of fishing
location choice. For this reason, the fleet dynamics sub-model
include\replaced[id=PD]{d}{s} a learning component, where a vessel's initial
fishing location in a trip \replaced[id=PD]{wa}{i}s based on selecting from
previously successful fishing locations. This \replaced[id=PD]{wa}{i}s achieved
by \added[id=PD]{calculating an expected profit based on the profit from
	locations previously fished} in the preceding trip as well as the same
month periods in preceding years, and choosing randomly from the top 75 \% of
fishing events \replaced[id=PD]{as defined by the expected profit}{in value}.
\added[id=PD]{Expected profit was estimated from the revenue achieved in
	previous fishing events at a location minus the fuel cost of travelling
	from the currently location to the new location.} Simulation testing
indicated that this learning increased the mean value of catches for the
vessels, over just relying on the correlated random walk
\todo{\added[id=JJ]{Correlated random walk of what}} function \added[id=PD]{as
	described for the 'within trip' decisions below} (MIGHT NEED TO INCLUDE
IN SUPPLEMENTARY).

@@

We parametrised the simulation model for four populations with differing
habitat preference\replaced{,}{ and} temperature tolerances (Figures
\ref{S-fig:1}, \ref{S-fig:3}, \ref{S-fig:4}, \ref{S-fig:5}, \ref{S-fig:6},
\ref{S-fig:7}), population demographic and recruitment functions. In addition,
each of the populations has two defined spawning areas which result in the
populations moving towards these areas in given weeks (Figure \ref{S-fig:2})
and population-specific movement rates (Table \ref{tab:1}). The realised
movement of the populations for a number of weeks is shown in Figure
\ref{S-fig:9} while the realised daily fishing mortality are shown in Figure
\ref{S-fig:10}.

>>>>

\usepackage{lineno,hyperref}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows,chains,calc}
\usepackage{hhline}
\usepackage{booktabs}
\usepackage[textsize=scriptsize]{todonotes}
%\usepackage[textsize=scriptsize, disable]{todonotes}  %% to turn off notes
\usepackage{eurosym}
\usepackage{longtable}

@@

We simulate 50 years of fishing and use the results from the fisheries catch to
draw inference on the underlying population structures. We compare this
inference to i) a simulated fixed-site sampling design commonly used for
fisheries monitoring purposes, and ii) the true underlying population
structures input to the simulation. We use the results to establish the
potential and limitations of fishery-dependent data in providing a robust
picture of spatiotemporal distributions. Finally, we simulate an area closure
based on areas defined from the known ("real-population") distribution,
commercial catch data and survey data at different temporal and spatial
resolutions and assess their effectiveness on reducing catches of a fish
population. \\

We conclude from our simulations that commercial data, while containing bias,
provides a useful tool for managing catches in mixed fisheries if applied at
the correct spatiotemporal scale. \\

@@

Fishers exploit a variety of fish populations that are heterogeneously
distributed in space and time with varying knowledge of species distributions
using species-unselective fishing gear. \added[id=PD]{In doing so} fisheries
\deleted[id=PD]{that}catch an assemblage of species
\added[id=PD]{and}\deleted{, known as mixed fisheries} \replaced[id=JJ]{may}{,
	when managed by single-species quotas can end up}
discard\deleted[id=JJ]{ing} overquota catch \added[id=JJ]{when managed by
	single species quotas and fishers exhaust or more quota, may} lead to
overexploitation of fish populations \added[id=JJ]{\citep{Ulrich2011a,
		Batsleer2015}}.  \added[id=JJ]{This discarding of fish in
	excess of quota hampers the ability to limit fishing mortality to
	within sustainable limits \citep{Alverson1994, Crowder1998,
		Rijnsdorp2007}} \deleted[id=PD]{; reducing discarding is
	crucial}and the ability to manage for the biological and economic
sustainability of fisheries\deleted[id=JJ]{ and implementation of an ecosystem
	approach to fisheries}. As such, there is increasing interest in
technical solutions such as gear and spatial closures as ways of
\replaced[id=PD]{reducing unwanted catch}{avoiding discard\added[id=JJ]{ing of
		fish}}\citep{Kennelly2002, Catchpole2008, Bellido2011}.  \\

\replaced[id=PD]{Changes to spatial fishing patterns have}{Use of spatial
	management as a tools} been proposed as a method to reduce discards
\added[id=PD]{\citep{Holmes2011, Little2014, Dunn2014a}}. However,
\deleted[id=PD]{its} implementation of avoidance measures is hampered by lack
of knowledge of fish and fishery spatiotemporal dynamics and understanding of
the scale at which processes are important for management. Understanding the
correct scale for spatial measures is crucial in order to implement measures at
a resolution that ensures effective management \citep{Dunn2016} while
minimising economic impact.  For example, a scale that promotes species
avoidance for vulnerable or low quota species while allowing continuance of
sustainable fisheries for available quota species.\\

\replaced[id=PD]{Identifying}{Ensuring measures are implemented at} an
appropriate scale has been a challenge in the past that has led to ineffectual
measures with unintended consequences such as limited impact towards the
management objective or increased benthic impact on previously unexploited
areas (e.g. the cod closure in the North Sea
\citep{Rijnsdorp2001,Dinmore2003}). \replaced[id=PD]{M}{Since then m}ore
refined spatial information has \added[id=PD]{since} become available through
the combination of logbook and Vessel Monitoring System (VMS) data
\citep{Lee2010, Bastardie2010, Gerritsen2012, Mateo2016} and more real-time
spatial management has been possible \citep[e.g.][]{Holmes2011}.  However, such
information is derived from an inherently biased sampling programme, targeted
fishing. \deleted[id=PD]{ Further, fishers generally only recorded landings
	(not catch) on a daily basis\todo{\added[id=JJ]{Its not clear what the
			problem is: landings collected on daily basis or
			landings recorded rather than catch}}.  This leads to
	questions about the validity of inference that can be drawn from
	landings data assigned to VMS activity pings.}\todo{\added[id=JJ]{This
		comes as a surprise: I thought this was going to be about
		discards}\added[id=PD]{Agree, have removed this to avoid
		confusion}}\\

@@

\begin{enumerate}
	\item to understand how sampling-derived data reflects the underlying
		population structures. We compare at different spatial and
		temporal aggregations of the simulated population distributions
		to: 
		\begin{enumerate}
			\item the inferred population from a stratified
				fixed-site sampling survey design commonly used
				for fisheries monitoring purposes, otherwise
				know as a fisheries-independent survey,
			\item the inferred population from our
				fishery-dependent model which includes
				fishery-induced sampling dynamics.\\
		\end{enumerate}
	\item to understand the impact of data aggregation and data source on
		spatial fisheries management measures we simulate a fishery
		closure to protect a species based on different spatial and
		temporal data aggregations:
		\begin{enumerate}
			\item as if the real spatial population structure were
				known,
			\item the fishery-independent inferred population
				structure
			\item the fishery-dependent inferred population
				structure
		\end{enumerate}
		We evaluate the theoretical "benefit" to the population of the
		closure(s), the effect on the other three populations and
		fishery catch.\todo{\added[id=JJ]{If the paper has two goals
				this should be clear from the start, but may be
				better over 2 MSs}\added[id=PD]{I would like to
				keep both parts, but have made clearer in how
				its set out. The closure scenarios form
				validation of the data aggregation, rather than
				effectiveness of the closures themselves - so
				its a continuation of the same question in my
				eyes}}\deleted[id=PD]{.  Further, we extend our
			analysis to a range of spatial and temporal scales to
			assess the impact of these processes on the success of
			the management measure.}\\
\end{enumerate}

@@

Mortality $Z_{c,d}$ can be decomposed to natural mortality, $M_{c,d}$, and
fishing mortality, $F_{c,d}$, where both $M_{c,d}$ and $F_{c,d}$ are
instantaneous rates with $M_{c,d}$ fixed and $F_{c,d}$ calculated by solving
the Baranov catch equation \citep{Hilborn1992b} for $F_{c,d}$:
\begin{equation*}
C_{c,d} = \frac{F_{c,d}}{F_{c,d} + M_{c,d}} * (1 - e^{-(F_{c,d} + M_{c,d})}) *
B_{c,d}
\end{equation*}
where $C_{c,d}$ is the summed catch from the fishing model across all fleets
and vessels in cell $c$ for the population during the day $d$, and $B_{c,d}$
the daily biomass for the population in the cell. Here, catch and fishing
mortality are the sum of those across all fleets and vessels, where $F_{fl, v,
	c, d, p} = E_{fl, v, c, d} \cdot Q_{fl, p} \cdot B_{c, d, p}$ with
$fl$, $v$ and $p$ the fleet, vessel and population respectively and $E$ and $Q$
fishing effort and catchability. \todo{\added[id=CM]{[link $F$ to effort and
		catchability - as I think we have F as an emergent property of
		the fleets rather than something we solve for (I could be wrong
		though!) - catch for a vessel is a product of catchability and
		biomass, i.e. C = qB, but this catch is summed to solve for F.
		So its both really]}}\\

@@

Recruitment is modelled through a function relating the mature biomass to
recruits at time of recruitment. In \emph{MixFishSim}, it can be modelled
either either as a stochastic Beverton-Holt stock-recruit form
\citep{Beverton1957}: 
\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & \frac{(\alpha * B_{c,d})}{(\beta + B_{c,d})} \\
	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}
Where $\alpha$ is the maximum recruitment rate, $\beta$ the spawning stock
biomass (SSB) required to produce half the maximum, $B$ current SSB and
$\sigma^2$ the variability in the recruitment due to stochastic
processes, or a stochastic Ricker form \citep{Ricker1954}:
\begin{equation*}
	\begin{split}
	\bar{R}_{c,d} = & B_{c,d} * e^{(\alpha - \beta * B_{c,d})} \\	
   	     R_{c,d} \sim & \log N[(\log(\bar{R}_{c,d}),\log(\sigma^2))]
	\end{split}
\end{equation*}
where $\alpha$ is the maximum productivity per spawner and $\beta$ the density
dependent reduction in productivity as the SSB increases. In this study, the
Beverton-Holt form of stock recruit relationship was used for all populations.

@@

To simulate \deleted[id=JJ]{how} fish population\deleted[id=JJ]{s might be}
distribut\replaced[id=JJ]{ion}{ed} in space and time\deleted[id=JJ]{, we
	employed} a Gaussian spatial process \added[id=JJ]{was employed} to
model habitat suitability for each of the populations on a 2d
grid\replaced[id=JJ]{. An}{, with an} advection-diffusion process
\deleted[id=JJ]{to} control\added[id=JJ]{led} \deleted[id=JJ]{how the}
population\deleted[id=JJ]{s} move\replaced[id=JJ]{ment}{d}, with a time-varying
temperature covariate used to change the spatial bounds of suitable habitat
on a weekly time-step\todo{\added[id=JJ]{What have a temperature covariate? Could just
		use time}\added[id=PD]{Was intended as some biological meaning
		- species thermal tolerances load onto the temperature effect -
		so could be different per species}}.  \deleted[id=JJ]{This was
	intended to balance realism in population movement, capturing the main
	directed and random processes, and practicality of modelling the
	population rather than individual fish.} \\

For \deleted[id=PD]{the}habitat we first define\added[id=PD]{d} a Gaussian
random field process\todo{\added[id=JJ]{Not clear how habitat/GRF affect local
		abundances, only have $B_{y,d}$}\added[id=PD]{Have included
		cell reference, $c$ to make spatial link explicit}}, $\{S(c) :
c \in \mathbb{R}^2\}$, \deleted[id=PD]{that is a stochastic process} where
\added[id=PD]{for} any set of cells $c_{1}, \dots, c_{n}$ \deleted[id=PD]{where
	for each $c_{i} \in \mathbb{R}^2$}, the joint distribution of $S =
\{S(c1),\dots S(c_{n})\}$ is multivariate Gaussian. The distribution is
specified by its \textit{mean function}, $\mu(c) = E[S(c)]$ and its
\textit{covariance function}\todo{\added[id=JM]{Introduce the gamma function,
		and why this covariance structure? Why correlate values in the
		random field?}\added[id=PD]{to allow populations to have
		different aggregation densities: have tried to clarify}},
$\gamma(c,c') = Cov\{S(c),S(c')\}$ \citep{Diggle2007}.\\

The covariance structure affects the smoothness of the surfaces which the
process generates\replaced[id=PD]{;}{, and} we used the \textit{Matérn}
\deleted[id=PD]{family of}covariance structure\deleted[id=PD]{s},
\replaced[id=PD]{where}{one where} the correlation strength weakens with
distance\deleted[id=PD]{(i.e. the correlation between $S(x)$ and $S(x')$
	decreases as the distance $u = ||x - x'||$ increases)}.
\added[id=PD]{This enables us to model the spatial autocorrelation observed in
	animal populations where density is more similar in nearby locations
	\citep{Tobler1970, F.Dormann2007}} and we change the parameters to
implement different spatial structures for the populations. The \textit{Matérn}
correlation is a two-parameter family where: 
	\begin{center}
		$\rho(u) = \{2^{\kappa -
			1}\Gamma{\kappa}\}^{-1}(u/\phi)^{\kappa}K_{\kappa}(u/\phi)$
	\end{center}
$K_{\kappa}(.)$ is a modified Bessel function of order $\kappa$, $\phi >
0$ is a scale parameter with the dimensions of distance, and $\kappa > 0$,
called the order, is a shape parameter which determines the smoothness of the
underlying process (Figure \ref{S-fig:16}). \\

\replaced[id=PD]{T}{In the simulation model, t}he habitat for each of the
populations \replaced[id=PD]{was}{is} generated \replaced[id=PD]{with}{through}
the \textit{RFSimulate} function of the \textit{RandomFields} R package
\citep{Schlater2015}. Each population \replaced[id=PD]{was}{is} initialised at
a single location, and subsequently move\replaced[id=PD]{ed}{s} according to a
probabilistic distrib\added[id=PD]{u}tion based on habitat suitability
(represented by the normalised values from the GRFs), temperature and distance
from current cell\replaced[id=PD]{:}{.} 
\begin{equation}
	Pr(J | I) = \frac{e^{-\lambda * d_{IJ}} \cdot
		(Hab_{J,p}^2 \cdot Tol_{J,p,wk})}{\sum\limits_{c=1}^{C}e^{-\lambda * d} \cdot
		(Hab_{c,p}^2 \cdot Tol_{c,p,wk})}
\end{equation}
Where $d_{IJ}$ is the euclidean distance between cell $I$ and cell $J$,
$\lambda$ is a given rate of decay, $Hab_{J,p}^2$ is the squared index of
habitat suitability for cell $J$ and population $p$, with $Tol_{J,p,wk}$ the
temperature tolerance for cell $J$ by population $p$ in week $wk$.\\

During pre-defined weeks of the year the habitat quality is modified with
\added[id=PD]{user-defined} spawning habitat locations\deleted[id=PD]{s},
\replaced[id=PD]{resulting in}{meaning} each population
\replaced[id=PD]{having}{has} concentrated areas where spawning takes place. In
the simulations the populations move\replaced[id=PD]{d}{s} towards
\replaced[id=PD]{these cells}{this} in the weeks prior to spawning, resulting
in directional movement towards the spawning grounds.\todo{\added[id=JM]{What
		does it mean concisely? Areas are assigned?} \added[id=PD]{Yes,
		the areas are pre-defined - I have amended to reflect and tried
		to clarify}} \\

The temperature field \replaced[id=PD]{was}{is} defined on a gradient from a
South-Westerly to North-Easterly direction, with temperature in each cell
changing gradually on a week-by-week basis so that initially high temperature
areas cycle\added[id=PD]{d} to lower temperatures and low temperature areas
\textit{vice versa}. Each population $p$ \replaced[id=PD]{wa}{i}s assigned a thermal
tolerance with mean, $\mu\added[id=PD]{_{p}}$ and variance,
$\sigma^2\added[id=PD]{_{p}}$ so that each cell and population temperature
suitability is defined that:
\begin{equation}
	Tol_{c, p\added[id=PD]{, wk}} = \frac{1}{ \sqrt (2\pi \cdot \sigma^2_{p})} \cdot \exp(-
		\frac{(T_{c\added[id=PD]{, wk}} - \mu_{p})^2}{2 \cdot \sigma^2_{p}} )	
\end{equation}
Where $Tol_{c, p, \added[id=PD]{wk}}$ is the tolerance of population $p$ for
cell $c$ \added[id=PD]{in week $wk$}, $T_{c, \added[id=PD]{wk}}$ is the
temperature in the cell \added[id=PD]{given the week} and
$\mu\added[id=PD]{_{p}}$ and $\sigma^2\added[id=PD]{_{p}}$ the mean and
standard deviation of the population temperature tolerance. \\

\added[id=PD]{The final process resulted in independent populations structure
	and movement patterns, with population movement occurring on a weekly
	basis. This process approximated the demographic shifts in fish
	populations throughout a year with seasonal spawning patterns (e.g.
	Figure \ref{S-fig:5}).}

@@

The fleet dynamics can be broadly categorised into three components; fleet
targeting - which determine\replaced[id=PD]{d}{s} the fleet catch efficiency
and preference towards a particular species; trip-level decisions, which
determine\added[id=PD]{d} the initial location to be fished at the beginning of
a trip; and within-trip decisions, determining movement from one fishing spot
to another within a trip. Together, these element implement an explore-exploit
type strategy for individual vessels to maximise their catch from an unknown
resource distribution (\cite{Bailey2018}).

@@

Each fleet of \textit{n} vessels \replaced[id=PD]{was}{is} characterised by
both a general efficiency, $Q_{fl}$, and a population specific efficiency,
${Q_{fl, p}}$.  Thus, the product of these parameters [$Q_{fl} \cdot Q_{fl,
	p}$] affect\replaced[id=PD]{s}{ed} the overall catch rates for the
fleet and the preferential targeting of one population over another. This, in
combination with the parameter choice for the step-function
\added[id=PD]{defined below} (as well as some randomness from the exploratory
fishing process) determine\replaced[id=PD]{d}{s} the preference of fishing
locations for the fleet.  All species prices \replaced[id=PD]{were}{are} kept
the same across fleets \replaced[id=PD]{and seasons}{, though can be made to
	vary seasonally}.

@@

Several studies \citep[e.g.][]{Hutton2004, Tidd2012, Girardin2015} have
confirmed past activity and past catch rates are strong predictors of fishing
location choice. For this reason, the fleet dynamics sub-model
include\replaced[id=PD]{d}{s} a learning component, where a vessel's initial
fishing location in a trip \replaced[id=PD]{wa}{i}s based on selecting from
previously successful fishing locations. This \replaced[id=PD]{wa}{i}s achieved
by \added[id=PD]{calculating an expected revenue based on the catches from
	locations fished} in the preceding trip as well as the same month
periods in previous years and the travel costs from the port to the fishing
grounds, and choosing randomly from the top 75 \% of fishing events
\replaced[id=PD]{as defined by the expected profit}{in value}. Simulation
testing indicated that this learning increased the mean value of catches for
the vessels, over just relying on the correlated random walk
\todo{\added[id=JJ]{Correlated random walk of what}} function \added[id=PD]{as
	described for the 'within trip' decisions below} (MIGHT NEED TO INCLUDE
IN SUPPLEMENTARY).

@@

We parametrised the simulation model for four populations with differing
habitat preference\replaced{,}{ and} temperature tolerances (Figures
\ref{S-fig:1}, \ref{S-fig:3}, \ref{S-fig:4}, \ref{S-fig:5}, \ref{S-fig:6},
\ref{S-fig:7}), population demographics and recruitment functions. In addition,
each of the populations has two defined spawning areas which result in the
populations moving towards these areas in given weeks (Figure \ref{S-fig:2})
and population-specific movement rates (Table \ref{tab:1}). The realised
movement of the populations for a number of weeks is shown in Figure
\ref{S-fig:9} while the realised daily fishing mortality are shown in Figure
\ref{S-fig:10}.
